---
title: "Grade analysis lab"
author: "Edoardo Ruvolo"
format: html
echo: false
editor: visual
---

```{r}
#| message: false
here::i_am("r-101-grade-analysis.Rproj")
library(here)
library(vroom) # or readr
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
```

## Exercise 1

Github link 

### Question 1

For this first question we were asked to load the data set using a local file name (using here).

```{r}
#| message: false
grades <- vroom(here("/Users/edoardoruvolo/Desktop/Data Management/r-101-grade-analysis/grades.csv"))
```

### Question 2

For the second question, we computed the minimum, maximum, median and mean grade at the Exam and included the results as a table in the rendered document.

```{r}
MMMG <- grades %>%
  summarise('Min'=min(Exam, na.rm=TRUE), 
            'Max'=max(Exam, na.rm=TRUE), 
            'Mean'=mean(Exam, na.rm=TRUE), 
            'Median'=median(Exam, na.rm=TRUE))%>%
 knitr::kable()
MMMG
```

### Question 3

In this third question, we were asked to count the number of students who did not take the final exam (i.e. for whom Exam is NA) and include the results in a sentence in your rendered document.

```{r}
NNEX <- sum(is.na(grades$Exam))
```

On the day of the exam, `r NNEX` people did not participate to the exam

### Question 4

In question 4 we represented graphically in an adapted way the distribution of grades at the Exam.

```{r}
NG <- grades %>% 
  drop_na(Exam)
ggplot(NG, aes(x = Exam))+ 
  geom_bar(fill = "skyblue", color = "black") +
  xlab("Distribution of the exam grade")+
  ylab("Count")+
  ggtitle("Distribution of grades")
```

### Question 5

In question 5, we were asked to compute the number of students in each Group and include the results in your document as a table.

```{r}
BG <- grades %>%
  group_by(Group) %>%
  summarise(N=n())%>%
  arrange(desc(N)) 
BG %>%
  knitr::kable()
```

### Question 6

In question 6, we represented the number of student in each group in a graphical way.

```{r}
ggplot(BG, aes(x = Group, y = N))+ 
  geom_col(fill = "skyblue", color = "black", size = 0.5) +
  xlab("Group Name")+
  ylab("Count of students")+
  ggtitle("Number of student per group")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Question 7

In question 7 we were asked to represent graphically the distribution of the grades at the Exam, by conditioning it a the group. We needed to propose two different graphical representation.

In the question 7, a bar graph is proposed, which in one solution is able to describe the distribution of the grade of the exam. In question 7.1 a facet wrap solution is proposed, showing the distribution of each group in a separated graph.

```{r}
GBG<- grades %>%
  filter(!is.na(Exam))
ggplot(GBG, aes(x = Exam, fill=Group)) + 
  geom_bar(position="fill") +
  ylab("Number of students in percentage")+
  xlab("Exam Score")+
 ggtitle("Distribution of Exam Scores by Group")
```

### Question 7.1

```{r}
ggplot(GBG, aes(x = Exam)) +
geom_bar(fill = "skyblue", color = "black", size = 0.5) +
  facet_wrap(~Group, scales = "free_y") +
  ylab("count")+
  xlab("Exam grade")+
  ggtitle("Distribution of Exam Scores by Group")
```

### Question 8

In question 8, we were asked to find the number of students who missed the exam in each group. A table with the first 4 values is presented just as a small sample.

```{r}
NG2 <- grades %>%
group_by(Group) %>%
summarise(not_attending = sum(is.na(Exam)))
table_NG2 <- knitr::kable(head(NG2, n= 4), caption = "Students not attending the exam in each group")
table_NG2
```

### Question 9

In question 9 we were asked to graphically represent the the number of students who missed the exam in each group. 
```{r}
ggplot(NG2, aes(x = Group, y = not_attending)) +
  geom_col(fill = "green", col = "darkgreen") +
  xlab("Groups")+
  ylab("Count of absent students")+
 ggtitle("Number of Absent Students by Group")+
 theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### Question 10

In question 10 we created a new data frame, reshaping the original data set to a long format. 
The new data frame should keep the Id and the Group as the original variables.

```{r}
GL <- grades %>%
  select(starts_with("MCQ"), starts_with("Online"), Exam, Id, Group) %>%
  pivot_longer(cols = starts_with("MCQ") | starts_with("Online") | Exam, names_to = "Examination", values_to = "Grade")
```

### Question 11

In question 11, using the long format, we computed the number of missing grades in total for each student. A table with the first 4 values is presented just as a small sample.

```{r}
MGS <- GL %>%
  group_by(Id) %>%
  summarise(Missing_Grades = sum(is.na(Grade)))
table_MGS <- knitr::kable(head(MGS, n = 4), caption = "Count of missing grade per student")
table_MGS
```

### Question 12

In question 12, we presented graphically the distribution of the number of missing grades per student.
```{r}
ggplot(MGS, aes(x = Missing_Grades)) +
  ylab("# of students") +
  xlab("# of missing grades") +
  ggtitle("Distribution of missing grades per student")+
  geom_histogram(color = "red", fill = "darkred", bins = 30)
```

### Question 13

In question 13 , using the long format, we found the number of students who missed the exam in each group
A table with the first 4 values is presented just as a small sample.
```{r}
NG3 <- GL %>%
  group_by(Id) %>%
  summarise(Absent = sum(is.na(Grade)))
table_NG3 <- knitr::kable(head (NG3, n=4) , caption = "Count of student")
table_NG3
```

### Question 14

In question 14, using the long format, we were asked to compute the number of missing grades for the online tests for each student. 
A table with the first 4 values is presented just as a small sample.
```{r}
MGO <- GL %>%
  filter(str_starts(Examination, "Online_MCQ")) %>%
  group_by(Id) %>%
  summarise(Missing_grade_online = sum(is.na(Grade)))
  table_MGO <- knitr::kable(head (MGO, n=4), caption = "Missing grade online per each student (ID)")
table_MGO
```

### Question 15

In question 15 we were asked to epresent graphically the distribution of the grades at the Exam conditioned by the number of missing grades for the online tests.
```{r}
Merged_DF1 <- grades %>%
   left_join(MGO, by = "Id")

DT1 = filter(Merged_DF1, !is.na(Exam))

ggplot(DT1, aes(x = as.factor(Missing_grade_online), y = Exam)) +
  geom_boxplot() +
  xlab("Total of missing grades for the online tests") +
  ylab("Grade at the exam") +
  ggtitle("Distribution of the grades at the exam, conditioned on the number of missing grades for the online tests")
```

### Question 16

In question 16 we created a table with two columns: Id to identify students and Missed with value TRUE when the student miss at least one MCQ_xx grade and FALSE when they miss no grade. A table with the first 10 values is presented just as a sample but in general the observation would be 729 (the Id of each student). 
```{r}
MMCQ <- grades %>%
  rowwise() %>%
  mutate(Missed = any(is.na(c_across(starts_with("MCQ"))))) %>%
  select(Id, Missed)
MMCQ <- MMCQ[order(MMCQ$Id, decreasing = FALSE),]
table_MMCQ <- knitr::kable(head (MMCQ, n=10), caption = "At least one MCQ missed, per student")
table_MMCQ
```

### Question 17

In question 17 we were asked to create a table with two columns: Group to identify groups and P_missed with the percentage of students in each group who missed at least one MCQ_xx grade. A table with the first 10 values is presented just as a sample but in general the observation would be 24 (each group). 
```{r}
PMMCQ <- grades %>%
  rowwise() %>%
  mutate(Missed = any(is.na(c_across(starts_with("MCQ"))))) %>%
  ungroup() %>%
  group_by(Group) %>%
  summarise(
    Total_student_per_group = n(),
    Students_missing_one = sum(Missed, na.rm = TRUE),
    P_missed = (Students_missing_one / Total_student_per_group ) * 100
  ) %>%
   mutate(P_missed = scales::percent(P_missed, scale = 1)) %>%
  select(Group, P_missed)
table_PMMCQ <- knitr::kable(head(PMMCQ, n=10), caption = "Percentage of students in each group who missed at least one MCQ")
table_PMMCQ
```

### Question 18
Lastly, in question 18, we were demanded to represent graphically the average grade at the Exam per group as a function of the percentage of missed MCQ_xx grade as defined in the previous question.
```{r}
AGG <- grades %>%
  group_by(Group) %>%
  summarise(Total_student_per_group = n(),
            Grade_of_Group = sum(Exam, na.rm = TRUE),
            Average_grade = Grade_of_Group /Total_student_per_group) %>%
  select(Group, Average_grade)

JT <- inner_join(AGG, PMMCQ)

ggplot(JT, aes(x = P_missed, y = Average_grade, color = Group)) +
  geom_point() +
  xlab("Percentage of missed MCQ grade") +
  ylab("Average grade at the exam, per group") +
  ggtitle("Average grade at the exam, per group, as a function of the percentage of missed MCQ") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
