---
title: "Grade_analysislab"
format: html
echo: false
editor: visual
---

```{r}
here::i_am("r-101-grade-analysis.Rproj")
library(here)
library(vroom) # or readr
library(dplyr)
library(tidyr)
library(ggplot2)
```

## Exercise 1
### Question 1
```{r}
grades <- vroom(here("/Users/edoardoruvolo/Desktop/Data Management/r-101-grade-analysis/grades.csv"))
```

### Question 2
```{r}
MMMG <- grades %>%
  summarise('Min'=min(Exam, na.rm=TRUE), 
            'Max'=max(Exam, na.rm=TRUE), 
            'Mean'=mean(Exam, na.rm=TRUE), 
            'Median'=median(Exam, na.rm=TRUE))%>%
  knitr::kable()
MMMG
```
### Question 3 
```{r}
nnex <- sum(is.na(grades$Exam))
```

On the day of the exam, `r nnex` people did not partecipate to the exam

### Question 4
Represent graphically in an adapted way the distribution of grades at the Exam. Make sure to adapt the code to avoid any error or warning from ggplot2.

```{r}
nograde <-grades %>% 
   filter(is.na(Exam)) %>%
   summarise(N=n())
```

```{r}
nograde <- grades %>% 
  drop_na(Exam)
  
ggplot(nograde, aes(x = Exam))+ 
  geom_bar() +
  xlab("Distribution of the exam grade")+
  ylab("Count")+
  ggtitle("Distribution of grades")
```
### Question 5 
Compute the number of students in each Group and include the results in your document as a table.
```{r}
bygroup <- grades %>%
  group_by(Group) %>%
  summarise(N=n()) %>%
  knitr::kable()
bygroup
```

